---
title: "AvrPm8_expression_analysis"
author: "Marion Müller"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introductoin

This document contains the code (bash & R) to conduct expression analysis of AvrPm8 in different powdery mildew strains

## Data

CDS of ISR_7

## Software 

R 
salmon
edgeR

## Salmon index

```
salmon index -t $BLUMERIA/00.1.2.ISR_7/Bgt_ISR7_CDS_v1_4_1.fasta -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1
```

## Run salmon quant on all samples

###  ISR_7 on Chinese Spring 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/Isr7_A_1.fq.gz -2 $RNASEQRAWBLUMERIA/Isr7_A_2.fq.gz -p 4 -o ISR_7_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/Isr7_B_1.fq.gz -2 $RNASEQRAWBLUMERIA/Isr7_B_2.fq.gz -p 4 -o ISR_7_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/Isr7_C_1.fq.gz -2 $RNASEQRAWBLUMERIA/Isr7_C_2.fq.gz -p 4 -o ISR_7_rep3 ;
```

### GBR_JIW2 on Chinese Spring 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410414.fastq.gz  -p 4 -o GBR_JIW2_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410415.fastq.gz  -p 4 -o GBR_JIW2_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410416.fastq.gz  -p 4 -o GBR_JIW2_rep3 ;
```

### CHE_96224 on Chinese Spring 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410408.fastq.gz  -p 4 -o CHE_96224_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410409.fastq.gz  -p 4 -o CHE_96224_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410410.fastq.gz  -p 4 -o CHE_96224_rep3 ;
```


### CHE_94202 on Chinese Spring 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410411.fastq.gz  -p 4 -o CHE_94202_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410412.fastq.gz  -p 4 -o CHE_94202_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410413.fastq.gz  -p 4 -o CHE_94202_rep3 ;
```

### S-1391 on Matador 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410417.fastq.gz  -p 4 -o S_1391_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410418.fastq.gz  -p 4 -o S_1391_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410419.fastq.gz  -p 4 -o S_1391_rep3 ;
```
### S-1459 on Matador 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410420.fastq.gz  -p 4 -o S_1459_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410421.fastq.gz  -p 4 -o S_1459_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR6410422.fastq.gz  -p 4 -o S_1459_rep3 ;
```

### THUN-12 on Timbo 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517426.fastq.gz  -p 4 -o THUN12_Timbo_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517427.fastq.gz  -p 4 -o THUN12_Timbo_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517428.fastq.gz  -p 4 -o THUN12_Timbo_rep3 ;
```

### T3-8 on Timbo 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517435.fastq.gz  -p 4 -o T3_8_Timbo_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517436.fastq.gz  -p 4 -o T3_8_Timbo_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -r $RNASEQRAWBLUMERIA/SRR2517437.fastq.gz  -p 4 -o T3_8_Timbo_rep3 ;
```

### CHN_17-40 on Chinese Spring 2dpi

```
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/D_01_1.fq.gz -2 $RNASEQRAWBLUMERIA/D_01_2.fq.gz -p 4 -o CHN_17_40_rep1 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/D_02_1.fq.gz -2 $RNASEQRAWBLUMERIA/D_02_2.fq.gz -p 4 -o CHN_17_40_rep2 ;
salmon quant -i $SALMONINDEX/Index_Bgt_ISR7_CDS_v1_4_1 -l A -1 $RNASEQRAWBLUMERIA/D_03_1.fq.gz -2 $RNASEQRAWBLUMERIA/D_03_2.fq.gz -p 4 -o CHN_17_40_rep3 ;
```

## Relative quantification

```
for i in */ ; do cd $i ; cut -f5 quant.sf > count ; cd .. ; done
paste */count | grep -v NumReads > AllCount ; 
ls   | grep rep | sed -z s'/\n/\t/g' > Names ; %manually remove space after last entry
cat Names AllCount > CountTable ;
cut -f1,2 CHN_17_40_rep1/quant.sf  > GeneInfo ;
paste GeneInfo CountTable > RNAseqSalmonCountonISR7.txt ;
```

## EdgeR analysis

```{r libraries, echo=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(edgeR)
library(reshape2)
```

### Load the data

```{r dataload}
counts<-read.table("RNAseqSalmonCountonISR7.txt",h=T, row.names=1)
head(counts)
count_sub <- counts[,c(2:28)]
head(count_sub )
gene_length <- counts[,c(1)]
gene_length
d <- DGEList(counts=count_sub,genes = gene_length)
d <- calcNormFactors(d, method="TMM")
rpkm<-rpkm(d,gene_length)
head(rpkm)
rpkm<-as.data.frame(rpkm)
head(rpkm)
```


### AvrPm8

Commands to select the AvrPm8 gene from the table:
```{r}
rpkm$Name<-row.names(rpkm)
AvrPm8all<-rpkm %>% filter(Name=="BgISR7-10067")
AvrPm8<-melt(AvrPm8all) %>% separate(variable, c('Isolate', 'Rep'),sep="_rep")
AvrPm8
```


Plot AvrPm8 rpkm values for all publish isolates isolates. For THUN-12 and T3-8, the replicates on cultivar Timbo were selected
```{r}
AvrPm8allpub<-AvrPm8all %>% select(ISR_7_rep1,ISR_7_rep2,ISR_7_rep3,CHE_96224_rep1,CHE_96224_rep2,CHE_96224_rep3,CHE_94202_rep1,CHE_94202_rep2,CHE_94202_rep3,GBR_JIW2_rep1,GBR_JIW2_rep2,GBR_JIW2_rep3,CHN_17_40_rep1,CHN_17_40_rep2,CHN_17_40_rep3,THUN12_Timbo_rep1,THUN12_Timbo_rep2,THUN12_Timbo_rep3,T3_8_Timbo_rep1,T3_8_Timbo_rep2,T3_8_Timbo_rep3,S_1459_rep1,S_1459_rep2,S_1459_rep3,S_1391_rep1,S_1391_rep2,S_1391_rep3)
AvrPm8pub<-melt(AvrPm8allpub) %>% separate(variable, c('Isolate', 'Rep'),sep="_rep")
AvrPm8pub
```


## Final Figure for publication

```{r, out.width="100%", echo=FALSE}
AvrPm8pub<-subset(AvrPm8pub, Isolate %in% c("ISR_7","CHE_96224","CHE_94202","CHN_17_40","THUN12_Timbo","T3_8_Timbo","S_1391","S_1459"))
ggplot2AvrPm8_publog2<-ggplot(AvrPm8pub, aes(x=factor(Isolate,c("ISR_7","CHE_96224","CHE_94202","CHN_17_40","THUN12_Timbo","T3_8_Timbo","S_1391","S_1459")),y=log2(value))) + geom_boxplot(fill="white") + geom_jitter(width=0.2,size=2,fill="gold1",color="black",shape=21,set.seed(3)) +ylab("rpkm") +xlab("Isolate") + ylim(0,12) + scale_x_discrete(labels=c("ISR_7","CHE_96224","CHE_94202","CHN_17_40","THUN-12","T3-8","S-1391","S-1459")) +
  theme_bw() + ylab("log2(rpkm)")  + theme(axis.text=element_text(size=12),axis.title=element_text(size=16))
ggplot2AvrPm8_publog2
pdf("AvrPm8_publishedlog2.pdf",width=10, height=4)
ggplot2AvrPm8_publog2
dev.off()
``` 

## GBR_JIW2

```{r, out.width="100%", echo=FALSE}
AvrPm8_JIW2<-subset(AvrPm8, Isolate %in% c("ISR_7","GBR_JIW2"))
AvrPm8_JIW2
ggplot2AvrPm8_JIW2publog2<-ggplot(AvrPm8_JIW2, aes(x=factor(Isolate,c("ISR_7","GBR_JIW2")),y=log2(value))) + geom_boxplot(fill="white") + geom_jitter(width=0.2,size=2,fill="gold1",color="black",shape=21,set.seed(3)) +ylab("rpkm") +xlab("Isolate") + ylim(0,12) + scale_x_discrete(labels=c("ISR_7","GBR_JIW2")) +
  theme_bw() + ylab("log2(rpkm)") + theme(axis.text=element_text(size=12),axis.title=element_text(size=16))
ggplot2AvrPm8_JIW2publog2
pdf("AvrPm8_JIW2publishedlog2.pdf",width=3, height=4)
ggplot2AvrPm8_JIW2publog2
dev.off()
``` 



`



